---
title: "Using cram_simulation() for CRAM with Known DGP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using cram_simulation() for CRAM with Known DGP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(cramR)
library(data.table)
library(DT)
```

## 🎯 What is `cram_simulation()`?

The `cram_simulation()` function performs **simultaneous policy learning and evaluation** under a **known data-generating process (DGP)**. It is useful for:

- Benchmarking CRAM on simulated datasets
- Measuring empirical **bias**, **variance**, and **confidence interval coverage**
- Supporting both synthetic (`dgp_X`) and empirical (`X`) covariate generation

---

## 📦 Inputs Overview

You must supply either:
- `X`: a dataset to bootstrap from (empirical DGP)  
**or**
- `dgp_X`: a function that simulates covariates  

You must also define:
- `dgp_D(X)`: treatment assignment function  
- `dgp_Y(D, X)`: outcome generation function  

---

## 📘 Example: Simulated Data with Binary, Discrete, and Continuous Covariates

```{r}
# Define data generation process (DGP) functions
X_data <- data.table::data.table(
  binary = rbinom(100, 1, 0.5),                 # Binary variable (0 or 1)
  discrete = sample(1:5, 100, replace = TRUE),  # Discrete variable (1 to 5)
  continuous = rnorm(100)                       # Continuous variable
)

dgp_D <- function(X) rbinom(nrow(X), 1, 0.5)

dgp_Y <- function(D, X) {
  theta <- ifelse(
    X[, binary] == 1 & X[, discrete] <= 2,  # Group 1: High benefit
    1,
    ifelse(X[, binary] == 0 & X[, discrete] >= 4,  # Group 3: High adverse effect
           -1,
           0.1)  # Group 2: Neutral effect
  )
  Y <- D * (theta + rnorm(length(D), mean = 0, sd = 1)) +
    (1 - D) * rnorm(length(D))  # Outcome for untreated
  return(Y)
}

# Parameters
nb_simulations <- 10
nb_simulations_truth <- 2
batch <- 5

# Perform CRAM simulation
result <- cram_simulation(
  X = X_data,
  dgp_D = dgp_D,
  dgp_Y = dgp_Y,
  batch = batch,
  nb_simulations = nb_simulations,
  nb_simulations_truth = nb_simulations_truth,
  sample_size = 50
)

# Access results
result$avg_delta_estimate
result$delta_empirical_bias
```

---

## 📊 Output Summary

```{r}
result$raw_results
```

```{r}
result$interactive_table
```

Returns a list containing:

- `raw_results`: A summary of key averaged metrics  
- `interactive_table`: An interactive HTML widget for quick exploration  

| Metric                             | Meaning                                      |
|------------------------------------|----------------------------------------------|
| Average Proportion Treated         | Share of samples treated by learned policy   |
| Average Delta Estimate             | Mean treatment effect (Δ) estimate           |
| Delta Empirical Bias               | Bias of Δ estimate against truth             |
| Delta Empirical Coverage           | CI coverage of Δ estimate                    |
| Average Policy Value Estimate      | Mean value of final policy                   |
| Policy Value Empirical Bias        | Bias against true policy value               |
| Policy Value Empirical Coverage    | CI coverage of policy value                  |

---

## 💡 Notes

- Uses **batch splitting** for honest policy learning
- Variance estimates use **influence-function-based asymptotics**
- Simulations are grouped by `sim_id` and averaged
- You can plug in `custom_fit` and `custom_predict` if needed

---

## 📚 See Also

- `cram_policy()` — for CRAM on real experimental/observational data  
- `cram_bandit_sim()` — for contextual bandits  
- `cram_ml()` — for general supervised or unsupervised ML  
