---
title: "Quick Start with CRAM"
# output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start with CRAM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(pkgdown.max_print = Inf, width = 1000)
library(cramR)
library(data.table)
library(glmnet)
library(caret)
```

## Introduction

The **Cram** package provides a unified framework for:

- 🧠 **Cram Policy (`cram_policy`)**: Learn and evaluate individualized binary treatment rules using Cram. Supports flexible models, including causal forests and custom learners. Common examples include whether to treat a patient, send a discount offer, or provide financial aid based on estimated benefit.

- 📈 **Cram ML (`cram_ml`)**: Learn and evaluate standard machine learning models using Cram. It estimates the expected loss at the population level, giving you a reliable measure of how well the final model is likely to generalize to new data. Supports flexible training via caret or custom learners, and allows evaluation with user-defined loss metrics. Ideal for classification, regression, and other predictive tasks.

- 🎰 **Cram Bandit (`cram_bandit`)**: Perform on-policy evaluation of contextual bandit algorithms using Cram. Supports both real data and simulation environments with built-in policies.

This vignette walks through these **three core modules**.

---

## Cram User file

For reproducible use cases, see the example script provided in the Cram GitHub repository:

[View `user_cram.R` on GitHub](https://github.com/yanisvdc/cramR/blob/master/inst/examples/user_cram.R)

---

## 1. `cram_policy()` — Binary Policy Learning & Evaluation

### Generate Simulated Data

```{r}
generate_data <- function(n) {
  X <- data.table(
    binary = rbinom(n, 1, 0.5),
    discrete = sample(1:5, n, replace = TRUE),
    continuous = rnorm(n)
  )
  D <- rbinom(n, 1, 0.5)
  treatment_effect <- ifelse(X$binary == 1 & X$discrete <= 2, 1,
                       ifelse(X$binary == 0 & X$discrete >= 4, -1, 0.1))
  Y <- D * (treatment_effect + rnorm(n)) + (1 - D) * rnorm(n)
  list(X = X, D = D, Y = Y)
}

set.seed(123)
data <- generate_data(1000)
X <- data$X; D <- data$D; Y <- data$Y
```

### Run `cram_policy()` with causal forest

```{r}
res <- cram_policy(
  X, D, Y,
  batch = 20,
  model_type = "causal_forest",
  learner_type = NULL,
  baseline_policy = as.list(rep(0, nrow(X))),
  alpha = 0.05
)
print(res)
```


### Case of categorical target Y

  Use `caret` and choose a classification method outputting probabilities i.e. using the key word `classProbs = TRUE` in `trainControl`, see the following as an example with a Random Forest Classifier:
  
  ```r
  model_params <- list(formula = Y ~ ., caret_params = list(method = "rf", trControl = trainControl(method = "none", classProbs = TRUE)))
  ```
  Also note that all data inputs needs to be of numeric types, hence for `Y` categorical, it should contain numeric values representing the class of each observation. No need to use the type `factor` for `cram_policy()`. 

### Custom Models with `cram_policy()`

Set `model_params` to `NULL` and specify `custom_fit` and `custom_predict`.

```{r}
custom_fit <- function(X, Y, D, n_folds = 5) {
  treated <- which(D == 1); control <- which(D == 0)
  m1 <- cv.glmnet(as.matrix(X[treated, ]), Y[treated], alpha = 0, nfolds = n_folds)
  m0 <- cv.glmnet(as.matrix(X[control, ]), Y[control], alpha = 0, nfolds = n_folds)
  tau1 <- predict(m1, as.matrix(X[control, ]), s = "lambda.min") - Y[control]
  tau0 <- Y[treated] - predict(m0, as.matrix(X[treated, ]), s = "lambda.min")
  tau <- c(tau0, tau1); X_all <- rbind(X[treated, ], X[control, ])
  final_model <- cv.glmnet(as.matrix(X_all), tau, alpha = 0)
  final_model
}

custom_predict <- function(model, X, D) {
  as.numeric(predict(model, as.matrix(X), s = "lambda.min") > 0)
}

res <- cram_policy(
  X, D, Y,
  batch = 20,
  model_type = NULL,
  custom_fit = custom_fit,
  custom_predict = custom_predict
)
print(res)
```

---

## 2. `cram_ml()` — ML Learning & Evaluation

### Regression with `cram_ml()`

Specify `formula` and `caret_params`conforming to the popular `caret::train()` and set an individual loss under `loss_name`.

```{r}
set.seed(42)
data_df <- data.frame(
  x1 = rnorm(100), x2 = rnorm(100), x3 = rnorm(100), Y = rnorm(100)
)

caret_params <- list(
  method = "lm",
  trControl = trainControl(method = "none")
)

res <- cram_ml(
  data = data_df,
  formula = Y ~ .,
  batch = 5,
  loss_name = "se",
  caret_params = caret_params
)
print(res)
```

### Classification with `cram_ml()`

All data inputs needs to be of numeric types, hence for `Y` categorical, it should contain numeric values representing the class of each observation. No need to use the type `factor` for `cram_ml()`. 

#### Case 1: Predicting Class labels

In this case, the model outputs hard predictions (labels, e.g. 0, 1, 2 etc.), and the metric used is **classification accuracy**—the proportion of correctly predicted labels.

- Use `loss_name = "accuracy"`
- Set `classProbs = FALSE` in `trainControl`
- Set `classify = TRUE` in `cram_ml()`


```{r, eval = requireNamespace("randomForest", quietly = TRUE)}
set.seed(42)

# Generate binary classification dataset
X_data <- data.frame(x1 = rnorm(100), x2 = rnorm(100), x3 = rnorm(100))
Y_data <- rbinom(nrow(X_data), 1, 0.5)
data_df <- data.frame(X_data, Y = Y_data)

# Define caret parameters: predict labels (default behavior)
caret_params_rf <- list(
  method = "rf",
  trControl = trainControl(method = "none")
)

# Run CRAM ML with accuracy as loss
result <- cram_ml(
  data = data_df,
  formula = Y ~ .,
  batch = 5,
  loss_name = "accuracy",
  caret_params = caret_params_rf,
  classify = TRUE
)

print(result)
```


#### Case 2: Predicting Class Probabilities

In this setup, the model outputs **class probabilities**, and the loss is evaluated using **logarithmic loss (`logloss`)**—a standard metric for probabilistic classification.

- Use `loss_name = "logloss"`
- Set `classProbs = TRUE` in `trainControl`
- Set `classify = TRUE` in `cram_ml()`

```{r, eval = requireNamespace("randomForest", quietly = TRUE)}
set.seed(42)

# Generate binary classification dataset
X_data <- data.frame(x1 = rnorm(100), x2 = rnorm(100), x3 = rnorm(100))
Y_data <- rbinom(nrow(X_data), 1, 0.5)
data_df <- data.frame(X_data, Y = Y_data)

# Define caret parameters for probability output
caret_params_rf_probs <- list(
  method = "rf",
  trControl = trainControl(method = "none", classProbs = TRUE)
)

# Run CRAM ML with logloss as the evaluation loss
result <- cram_ml(
  data = data_df,
  formula = Y ~ .,
  batch = 5,
  loss_name = "logloss",
  caret_params = caret_params_rf_probs,
  classify = TRUE
)

print(result)
```

In addition to using built-in learners via `caret`, `cram_ml()` also supports **fully custom model workflows**. You can specify your own:

- Model fitting function (`custom_fit`)
- Prediction function (`custom_predict`)
- Loss function (`custom_loss`)

See the vignette "Cram ML" for more details.

---

## 3. `cram_bandit()` — Contextual Bandits for On-policy Statistical Evaluation

Specify:

- `pi`: An array of shape **(T × B, T, K)** or **(T × B, T)**, where:

  - \( T \) is the number of learning steps (or policy updates)
  
  - \( B \) is the batch size
  
  - \( K \) is the number of arms
  
  - \( T \times B \) is the total number of contexts
  
  - In the natural 3D version, `pi[j, t, a]` gives the probability that the policy \( \hat{\pi}_t \) assigns arm `a` to context \( X_j \). In the 2D version, we only keep the probabilities assigned to the **chosen arm** \( A_j \) for each context \( X_j \) in the historical data - and not the probabilities for all of the arms \( a \) under each context \( X_j \).
  
- `arm`: A vector of length \( T \times B \) indicating which arm was selected in each context.

- `reward`: A vector of observed rewards of length \( T \times B \).

- `batch`: (optional) Integer batch size \( B \). Default is 1.

- `alpha`: Significance level for confidence intervals.


```{r}
set.seed(42)
T <- 100; K <- 4
pi <- array(runif(T * T * K, 0.1, 1), dim = c(T, T, K))
for (t in 1:T) for (j in 1:T) pi[j, t, ] <- pi[j, t, ] / sum(pi[j, t, ])
arm <- sample(1:K, T, replace = TRUE)
reward <- rnorm(T, 1, 0.5)

res <- cram_bandit(pi, arm, reward, batch=1, alpha=0.05)
print(res)
```

---

## Summary

- `cram_policy()`: Learn and evaluate a binary policy.
- `cram_ml()`: Learn and evaluate ML models.
- `cram_bandit()`: Cramming contextual bandits for on-policy statistical evaluation.

```{r cleanup-autograph, include=FALSE}
autograph_files <- list.files(tempdir(), pattern = "^__autograph_generated_file.*\\.py$", full.names = TRUE)
if (length(autograph_files) > 0) {
  try(unlink(autograph_files, recursive = TRUE, force = TRUE), silent = TRUE)
}

